\documentclass{cv}

\begin{document}
\name{Dennis Chen}
\sociallist{\social{\faLink}{\href{https://www.geometryexplorer.xyz}{geometryexplorer.xyz}}\sep\social{\Letter}{\href{mailto:proofprogram@gmail.com}{proofprogram@gmail.com}}}
\sociallist{\social{\faLinkedinSquare}{\href{https://www.linkedin.com/in/dennisrchen/}{dennisrchen}}\sep\social{\faGithub}{\href{https://github.com/chennisden}{chennisden}}}
\cvQuote{Improving yourself daily means increasing your self-worth. Turn adversity into strength!}{Yugami}

\cat{Education}

\act{Amador Valley High School}{Fall 2019}{present}

\desc{Graduating in 2023. Current unweighted GPA is 4.0, weighted GPA is 4.3.}

\act{Chabot Community College}{Fall 2020}{}

\desc{A in Linear Algebra. Dual enrollment.}

\cat{Work}

\act{AlphaStar Academy}{December 2020}{July 2021}

\desc{Developed math curriculum as a part-time employee.}

\cat{Volunteering and Programs}

%\act{Math Advance!}{?? 2021(?)}{present}
%
%\desc{President and founder of 501(c)(3) for math contests and classes.}

\act{Mathematical Advancement by Self Training}{March 2020}{present}

\desc{Current director and founder of free program for students preparing for high-level computational exams such as the American Invitational Mathematics Examination (AIME) and international equivalents. More details can be found on \href{https://mast.geometryexplorer.xyz}{mast.geometryexplorer.xyz}.}

\act{Youth EUCLID Association}{2017}{present}

\desc{Current Senior VP of Math Practice classes. Former Math Practice teacher, 2020-2021 VP of Math Practice classes, head of 2020 Summer AMC Class.}

\refer{Reference: Mr. Jie Li.}

\act{Everaise Academy}{June 2020}{August 2020}

\desc{Former Math I Associate Head, minor role in Math II for the 2020 Summer Camp.}

\cat{Selected awards}

\act{USA Junior Math Olympiad (USAJMO) qualifier}{2020}{}

\act{American Invitational Mathematics Examination qualifier}{2019, 2020, 2021}{}

\cat{Publications}

\act{Proofs in Competition Math: Volume 1}{}{}

\refer{Toller, Alex, et al. Proofs in Competition Math. Vol. 1, Amazon Publishing, 2019.}

\act{Proofs in Competition Math: Volume 2}{}{}

\refer{Toller, Alex, et al. Proofs in Competition Math. Vol. 2, Amazon Publishing, 2019.}

\cat{Projects}

%\act{bounce.cls}{}{}
%
%\desc{Open-source \LaTeX{} class for articles and books. GitHub repository \href{https://www.github.com/chennisden/bounce}{here}. Sample \href{https://www.github.com/chennisden/bounce/tree/main/examples/article.pdf}{article} and \href{https://www.github.com/chennisden/bounce/tree/main/examples/book.pdf}{book}.}

\act{lucky.cls}{}{}

\desc{Open-source \LaTeX{} class primarily created for developing math materials. Sample document available \href{https://www.geometryexplorer.xyz/texfiles/CV/lucky.pdf}{here}.}

\cat{Skills}

\act{Java, C++, \LaTeX, Asymptote, TikZ, HTML, CSS}{}{}
\end{document}